The Diminishing Returns of Music in the Digital Era

### 2024-04-18 paper/paper.qmd
final final changes

The RIAA dataset published by @RIAA was cleaned to account for the irrelevance of the "number_of_records" and "metric" columns. Additional variables such as "average_revenue_per_format" and "yearly_revenue_format" were created to help with the generation of certain graphs. They were created by adding up the revenue per format and either finding the average of it or grouping the revenue by year. Using the dataset, we were able to replicate all of the means and standard deviations of the music formats, as shown in @tbl-summary1. The summary statistics illustrate how the high standard deviations across the board reflect a wide variation in yearly revenue, which could be attributed to a certain format's lifecycle or other market changes over time. For example, 8-Track tapes, despite their irrelevance, have an intriguingly high mean revenue.

@@

music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

summary_stats <- music_data %>%
  group_by(format) %>%
  summarise(
    mean_revenue = mean(value_actual, na.rm = TRUE),
    sd_revenue = sd(value_actual, na.rm = TRUE)
  ) %>%
  ungroup()

kable(summary_stats, caption = "Summary Statistics for Music Formats")
```

@@

music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

yearly_format_revenue <- music_data %>%
  group_by(year, format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  ungroup()

ggplot(yearly_format_revenue, aes(x = year, y = total_revenue, fill = format)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(x = "Year", y = "Total Revenue ($)", fill = "Format") +
  theme(legend.position = "bottom")
```

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-averagerevenue
#| tbl-cap: "Music Industry Average Revenue per Format"
#| warning: false
#| results: asis
#| 
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

average_revenue_per_format <- music_data %>%
  mutate(value_actual = as.numeric(value_actual)) %>%
  filter(value_actual > 0) %>%
  group_by(format) %>%
  summarise(average_revenue = mean(value_actual, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(average_revenue))

@@

average_revenue_table <- kable(head(average_revenue_per_format, 20), 
                               padding = 5, 
                               format = "markdown",
                               digits = 2
                               )

@@

music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

top_formats <- music_data %>%
  group_by(format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  top_n(5, total_revenue) %>%
  pull(format)

@@

top_formats_data <- top_formats_data %>%
  group_by(format) %>%
  mutate(yoy_change = (value_actual / lag(value_actual) - 1) * 100) %>%
  ungroup()

ggplot(top_formats_data, aes(x = year, y = yoy_change, color = format)) +
  geom_line() +
  scale_y_continuous(limits = c(-100, 100),
                     labels = scales::percent) +
  labs(
       x = "Year",
       y = "Percentage Change",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

top_formats <- music_data %>%
  group_by(format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  top_n(7, total_revenue) %>%
  pull(format)

music_data_top <- music_data %>%
  filter(format %in% top_formats)

ggplot(music_data_top, aes(x = year, y = value_actual, color = format, group = format)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) + # Use scales package to format y axis
  labs(
       x = "Year",
       y = "Total Revenue",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

bayesian_model <- brm(value_actual ~ format, data = music_data, 
                      family = gaussian(), 
                      prior = c(set_prior("normal(0, 10)", class = "b"),
                                set_prior("cauchy(0, 2)", class = "sigma")),
                      chains = 4, iter = 2000, warmup = 500, 
                      control = list(adapt_delta = 0.95),
                      refresh = 0)
saveRDS(bayesian_model, file = "second_model.rds")
```

@@

modelsummary::modelsummary(
  list("Bayesian Linear Regression" = bayesian_model),
  output = "markdown",
  fmt = 2
)
```

>>>>

The RIAA dataset published by @RIAA was cleaned to account for the irrelevance of the "number_of_records" and "metric" columns. Additional variables such as "average_revenue_per_format", "top_formats", and "yearly_revenue_format" were created to help with the generation of certain graphs. They were created by adding up the revenue per format and either finding the average of it or grouping the revenue by year. Using the dataset, we were able to replicate all of the means and standard deviations of the music formats, as shown in @tbl-summary1. The summary statistics illustrate how the high standard deviations across the board reflect a wide variation in yearly revenue, which could be attributed to a certain format's lifecycle or other market changes over time. For example, 8-Track tapes, despite their irrelevance, have an intriguingly high mean revenue.

@@

# read the dataset from a parquet file
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

# develop summary statistics
summary_stats <- music_data %>%
  group_by(format) %>%
  summarise(
    mean_revenue = mean(value_actual, na.rm = TRUE),
    sd_revenue = sd(value_actual, na.rm = TRUE)
  ) %>%
  ungroup()

# display the summary statistics
kable(summary_stats, caption = "Summary Statistics for Music Formats")
```

@@

# read the dataset from a parquet file
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

# determine the yearly revenue per format
yearly_format_revenue <- music_data %>%
  group_by(year, format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  ungroup()

# plot the yearly revenue per format
ggplot(yearly_format_revenue, aes(x = year, y = total_revenue, fill = format)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(x = "Year", y = "Total Revenue ($)", fill = "Format") +
  theme(legend.position = "bottom")
```

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-averagerevenue
#| tbl-cap: "Music Industry Average Revenue per Format"
#| warning: false
#| results: asis

# read the dataset from a parquet file
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

# determine the average revenue per format
average_revenue_per_format <- music_data %>%
  mutate(value_actual = as.numeric(value_actual)) %>%
  filter(value_actual > 0) %>%
  group_by(format) %>%
  summarise(average_revenue = mean(value_actual, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(average_revenue))

@@

# create a table of the average revenue per format
average_revenue_table <- kable(head(average_revenue_per_format, 20), 
                               padding = 5, 
                               format = "markdown",
                               digits = 2
                               )

@@

# read the dataset from a parquet file
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

# find the top 5 music formats based on total revenue.
top_formats <- music_data %>%
  group_by(format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  top_n(5, total_revenue) %>%
  pull(format)

@@

# calculate the year-over-year percentage change in revenue for each format
top_formats_data <- top_formats_data %>%
  group_by(format) %>%
  mutate(yoy_change = (value_actual / lag(value_actual) - 1) * 100) %>%
  ungroup()

# plot the year-over-year percentage change in revenue for each format
ggplot(top_formats_data, aes(x = year, y = yoy_change, color = format)) +
  geom_line() +
  scale_y_continuous(limits = c(-100, 100),
                     labels = scales::percent) +
  labs(
       x = "Year",
       y = "Percentage Change",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

# read the dataset from a parquet file
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

# group the data by the "format" column, calculate the total revenue for each format and then select the top 7 formats by total revenue
top_formats <- music_data %>%
  group_by(format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  top_n(7, total_revenue) %>%
  pull(format)

# filter the original music_data to include only the rows where the format is one of the top formats
music_data_top <- music_data %>%
  filter(format %in% top_formats)

# plot the data using ggplot
ggplot(music_data_top, aes(x = year, y = value_actual, color = format, group = format)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) + # Use scales package to format y axis
  labs(
       x = "Year",
       y = "Total Revenue",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

# read the dataset from a parquet file
music_data <- read_parquet("/Users/victortimir/music-industry-analysis/data/analysis_data/musicdata.parquet")

# construct a bayesian model using the brm function from the rstanarm package
bayesian_model <- brm(value_actual ~ format, data = music_data, 
                      family = gaussian(), 
                      prior = c(set_prior("normal(0, 10)", class = "b"),
                                set_prior("cauchy(0, 2)", class = "sigma")),
                      chains = 4, iter = 2000, warmup = 500, 
                      control = list(adapt_delta = 0.95),
                      refresh = 0)

# save the bayesian model to an RDS file for future use
saveRDS(bayesian_model, file = "second_model.rds")
```

@@

# display the summary of the bayesian model in a table
modelsummary::modelsummary(
  list("Bayesian Linear Regression" = bayesian_model),
  output = "markdown",
  fmt = 2
)
```
